Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning

### 2021-08-28 2021/Entropy/entropy.tex
before submission

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
\usepackage{inputenc}
\usepackage[T2A,T1]{fontenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%

@@

For example, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
The solution to this class of problems, as a rule, cannot be found analytically and, therefore, one needs to construct numerical methods to solve it.

@@

There are several approaches to the construction of numerical methods for solving global optimization problems. 
%Надо ли добавить алгоритмы на основе метамоделей ?
Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters. Local optimization methods have a high convergence rate. At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions. 
Machine learning methods can be successfully applied to solve this problem.  
For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
In \cite{Cassioli2012}, the area for launching the local method was allocated based on the classification of starting points using a support vector machine.

@@

One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}. This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints and multicriteria optimization problems. For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.

@@

In this paper, we will consider global optimization problems of the form
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in \text{R}^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extreme, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.

@@

Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ и $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.

>>>>

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
%\usepackage{inputenc}
%\usepackage[T2A,T1]{fontenc}
%\usepackage[english,russian]{babel}
%\usepackage{cmap}
%%%%

@@

Particularly, ML could be applied for solving such a complex problem of computational mathematics as global optimization.
The solution to this class of problems, as a rule, cannot be found analytically and, therefore, one needs to construct numerical methods to solve it.

@@

There are several approaches to the construction of numerical methods for solving global optimization problems. 
%Надо ли добавить алгоритмы на основе метамоделей ?
Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters. Local optimization methods have a high convergence rate. At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions. 
Machine learning methods can be successfully applied to solve this problem.  
For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
In \cite{Cassioli2012} the area for starting the local method was allocated based on the classification of starting points using a support vector machine.

@@

One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}. This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints \cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}. For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.

@@

In this paper, we will consider global optimization problems of the form
\begin{eqnarray}\label{main_problem}
& \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}},\\
& D=\left\{y\in \text{R}^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}. \nonumber
\end{eqnarray}
Problem (\ref{main_problem}) is considered under the assumption that the objective function is multi-extremal, is given in a form of a ``black box'', and the calculation of its values is associated with solving the problem of numerical simulation, which makes the solution a labor-intensive operation.

@@

Tables \ref{table:average_GKLS01} and \ref{table:average_GKLS002} show the average number of trials required by the GSA and GSA-DT methods to correctly solve all problems with an accuracy of $\epsilon = 10^{-2}\left|b-a\right|$ and $\epsilon = 2 \cdot 10^{-3}\left|b-a\right|$ respectively.
